### 2020-10-27 doc/manuscript/Dow_manuscript.Rmd
structuring/ writing

#Figure numbers
schematic=1 #illustrates parameters
growth_pheno_by_xylem = 2  #if we include more sites, we'll make this multi-panel
rel_growth = "2a" #Based on D'Orangeville Figure 3.a), but with relative growth
cum_growth = "2b" #Similar to D'Orangeville Figure 4.a), but with cumulative relative growth
climwin = 3
pheno_Tsensitivity = 4
model_summary = 5
DOYtiming = 6
#SI:
#tables

#figures
climwin_rp_DOY25_SCBI = "S1"
climwin_dp_DOY25_SCBI = "S2"

#captions

schematic_caption <- "Figure 1. Schematic illustrating the parameters considered here. Shown are measurements for an example tree (## cm SPECIES), fit with the model of McMahon & Parker (2015), from which phenology and growth rate parameters are obtained. ... (HERE, WE NEED TO ANNOTATE WITH LABELS FOR EACH PARAMETER CONSIDERED IN THE MANUSCRIPT. I'D ALSO DROP THE BACKGROUND SHADING--MAKES IT BUSY.)"

rel_growth_caption <- "Figure 2a. (SHOULD BE MERGED INTO SINGLE IMAGE WITH 2B. SEE BELOW FOR CAPTION)"

cum_growth_caption  <- "Figure 2. Growth trajectories for ring- and diffuse-porous trees, as both relative (a) and cumulative (b) fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model."

climwin_caption <- "Figure 3. Landscapes of relationships between the day of year on which 25% of annual growth is achieved (DOY_25) and temperature in prior weeks for ring- and diffuse-porous trees. Shown is climwin output indicating the change in AIC relative to a null model (first panel) and linear coefficients (second panel). Window Open and Window Close indicate number of weeks prior to DOY_25 (ring-porous: DATE; diffuse-porous: DATE). Black boxes indicate the optimal window (i.e., that with lowest dAIC)."

@@

climwin_rp_DOY25_SCBI_short_caption <- "Figure S1. Full climwin output for DOY 25 for ring porous species at SCBI."
climwin_dp_DOY25_SCBI_short_caption <- "Figure S1. Full climwin output for DOY 25 for diffuse porous species at SCBI."
DOYtiming_caption <- "Figure 6. Day of year where growth milestones were achied at the Smithonian Conservation Biology Institute."
```

@@

#### (5. Here, we...)
Using many measurements taken within a year, we can model the growth of individual trees and extract data on growth milestones and traits for interspecific, intraspecific, and temporal comparison (McMahon & Parker, 2014). This allows the determination of variable responses of key tree species within a forest, meaning a better understanding of future growth under projected climate conditions. Here, we use biweekly dendrometer band measurements (5 years @ Harvard Forest, 10 @ SCBI) for dominant tree species at each site (15 @ Harvard Forest, 4 @ SCBI)  to characterize intraannual stem-growth phenology within two forest dynamics plots in the Eastern USA, calculating for each tree in each year where data was available: the day of year (DOY) where 25, 50, and 75% annual growth was achieved; the maximum growth rate; the DOY where maximum growth rate was achieved; the total growth; and season length (75%-25% DOY) (Fig. `r schematic`). Correlation between these variables and Spring temperatures over the 15 year study period (1998-2003; 2011-2019) were assessed using a linear mixed model in a heriarchal bayesian framework.

We test the hypotheses that (1) stem-growth milestones are reached earlier in years with warmer early springs, but (2) maximum growth rates are independent of spring temperatures, and as a result (3) total annual growth increases in response to warmer spring temperatures (Table `r hypothesis_table`).

@@

```{r hypothesis_table, eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)  
hypothesis_table <- read.csv(file = "tables_figures/hypothesis_table.csv")
knitr::kable(hypothesis_table, caption = "Table 1. Hypothesis table. (NEEDS FORMATTING...obviously!)")
```

@@

*Study sites and data*

@@

@@

(Bands used after minimum growth checks: 
1  2011    84
2  2012    84
3  2013   100
4  2014   100
5  2015    96
6  2016    94
7  2017    89
8  2018    88
9  2019    75) this seems like alot are being thrown out - manually check if WAI
Measurements were timed to begin before beginning of spring growth and to continue through the cessation of growth in the fall. *(Give specifics, which may vary by site. Note that we adjusted timing at SCBI when we observed early leaf-out of understory shrubs/herbs)* The median start date was 4/14, which was adjusted when early leaf-out of understory shrubs and herbs was observed, and the median end date was 10/17, coinciding with late senesence in the plot. The earliest start date for intraannual measurements was in 2020 on March 30, and the latest end date was on 2012 on November 26.

@@

*Fitting growth curves*

@@

*Determining optimal climate windows*

@@

@@

```{r climwin, out.width="100%", echo=FALSE, fig.align="center", fig.cap= cum_growth_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/climwin.png")
```

@@

@@

>>>>

#Figure numbers
schematic=1 #illustrates parameters
growth_pheno_by_xylem = 2  #if we include more sites, we'll make this multi-panel
rel_growth = "2a" #Based on D'Orangeville Figure 3.a), but with relative growth
cum_growth = "2b" #Similar to D'Orangeville Figure 4.a), but with cumulative relative growth
climwin = 3
pheno_Tsensitivity = 4
model_summary = 5
DOYtiming = 6
DOYtiming_SCBI = "6a"
DOYtiming_Harvard = "6b"

#captions
#tables
hypothesis_table_caption <-"Table 1. Hypothesis table. (NEEDS FORMATTING...obviously!)"

#figures
schematic_caption <- "Figure 1. Schematic illustrating the parameters considered here. Shown are measurements for an example tree (## cm SPECIES), fit with the model of McMahon & Parker (2015), from which phenology and growth rate parameters are obtained. ... (HERE, WE NEED TO ANNOTATE WITH LABELS FOR EACH PARAMETER CONSIDERED IN THE MANUSCRIPT. I'D ALSO DROP THE BACKGROUND SHADING--MAKES IT BUSY.)"

rel_growth_caption <- "Figure 2a. (SHOULD BE MERGED INTO SINGLE IMAGE WITH 2B. SEE BELOW FOR CAPTION)"
cum_growth_caption  <- "Figure 2. Growth trajectories for ring- and diffuse-porous trees, as both relative (a) and cumulative (b) fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model."

climwin_caption <- "Figure 3. Landscapes of relationships between the day of year on which 25% of annual growth is achieved (DOY_25) and temperature in prior weeks for ring- and diffuse-porous trees at SCBI and Harvard Forest. Shown are matrices of linear coefficients of first-order linear regressions between temperature and DOY_25, where Window Open and Window Close indicate number of weeks prior to DOY_25 (ring-porous: DATE at SCBI, DATE at HF; diffuse-porous: DATE at SCBI, DATE at HF). Black boxes indicate the optimal window (i.e., that with lowest dAIC).    (REPLOT ON THE SAME SCALE, WITH BLUE FOR POSITIVE AND RED FOR NEGATIVE)"

@@

DOYtiming_caption <- "Figure 6. Day of year where growth milestones were achied at (a) the Smithonian Conservation Biology Institute and (b) Harvard Forest."

# SUPPLEMENTARY INFO
#SI figure numbers
climwin_rp_DOY25_SCBI = "S1"
climwin_dp_DOY25_SCBI = "S2"
climwin_rp_DOY25_SCBI = "S3"
climwin_dp_DOY25_SCBI = "S4"
#SI figure captions
climwin_rp_DOY25_SCBI_short_caption <- "Figure S1. Full climwin output for DOY 25 for ring porous species at SCBI."
climwin_dp_DOY25_SCBI_short_caption <- "Figure S2. Full climwin output for DOY 25 for diffuse porous species at SCBI."
climwin_rp_DOY25_HARV_short_caption <- "Figure S3. Full climwin output for DOY 25 for ring porous species at Harvard Forest."
climwin_dp_DOY25_HARV_short_caption <- "Figure S4. Full climwin output for DOY 25 for diffuse porous species at Harvard Forest."

@@

#### (5. Here, we...)
Here, we characterize how early spring temperatures affect stem growth phenology and rates of temperate deciduous trees within **two/three** forest dynamics plots in the Eastern USA. Using data from dendrometer bands measured throughout the growing season, we fit a growth model  to the time series of individual trees to determine the day of year (DOY) where 25, 50, and 75% annual growth was achieved; maximum growth rates and the DOY on which they occurred; total annual growth; and peak growing season length (75%-25% DOY) [Fig. `r schematic`; @mcmahon_general_2015]. We test the hypotheses that (1) warmer early springs result in earlier stem growth and a period of growth, but (2) maximum growth rates are independent of spring temperatures, and as a result (3) total annual growth increases in response to warmer spring temperatures (Table `r hypothesis_table`).

@@

```{r hypothesis_table, eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)  
hypothesis_table <- read.csv(file = "tables_figures/hypothesis_table.csv")
knitr::kable(hypothesis_table, caption = hypothesis_table_caption)
```

@@

#### Study sites and data

@@

(5 years @ Harvard Forest, 10 @ SCBI)

@@

(Bands used after minimum growth checks: 
1  2011    84
2  2012    84
3  2013   100
4  2014   100
5  2015    96
6  2016    94
7  2017    89
8  2018    88
9  2019    75) this seems like alot are being thrown out - manually check if WAI

Measurements were timed to begin before beginning of spring growth and to continue through the cessation of growth in the fall. *(Give specifics, which may vary by site. Note that we adjusted timing at SCBI when we observed early leaf-out of understory shrubs/herbs)* The median start date was 4/14, which was adjusted when early leaf-out of understory shrubs and herbs was observed, and the median end date was 10/17, coinciding with late senesence in the plot. The earliest start date for intraannual measurements was in 2020 on March 30, and the latest end date was on 2012 on November 26.

@@

#### Fitting growth curves

@@

#### Determining optimal climate windows

@@

#### Hypothesis testing

Correlation between these variables and Spring temperatures over the 15 year study period (1998-2003; 2011-2019) were assessed using a linear mixed model in a heriarchal bayesian framework.

@@

```{r climwin, out.width="100%", echo=FALSE, fig.align="center", fig.cap= climwin_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/climwin.png")
```

@@

```{r DOYtiming2, out.width="100%", echo=FALSE, fig.align="center", fig.cap= DOYtiming_caption}
knitr::include_graphics("tables_figures/DOYtiming_HF_PRELIM.png")
```

@@

`r climwin_rp_DOY25_HARV_short_caption`

`r climwin_dp_DOY25_HARV_short_caption`
